Exploring the Traveling Salesman Problem

### 2017-05-02 paper/paper.tex
progress check

% Packages
% ---
\usepackage[
  left=3cm,
  right=3cm,
  top=2cm,
  bottom=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{clrscode3e}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{parskip}
\usepackage{mathptmx}
\usepackage{bigstrut}
\usepackage{mathptmx}
\usepackage{minted}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}

@@

% Document
% ---
\begin{document}

@@

\section{A More Thoughtful Solution}

% Appendicies
% ---
\newpage
\appendix
\section{Appendices}
\addcontentsline{toc}{section}{Appendices}
\renewcommand{\thesubsection}{\Alph{subsection}}

\subsection{Collaboration}
  I collaborated with Grant Eaton and Luke Artnak.

\newpage
\subsection{Code Demonstration}
  \inputminted[frame=lines,linenos=true]{python}{../src/brute_force.py}

>>>>

% Packages
% ---
\usepackage[
  left=3cm,
  right=3cm,
  top=2cm,
  bottom=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{clrscode3e}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{parskip}
\usepackage{mathptmx}
\usepackage{bigstrut}
\usepackage{mathptmx}
\usepackage{minted}
\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
\usepackage{apacite}
%\usepackage{natbib}

@@

% Document
% ---
\begin{document}
%\setlength{\abovedisplayskip}{0pt}
%\setlength{\belowdisplayskip}{0pt}
%\setlength{\abovedisplayshortskip}{0pt}
%\setlength{\belowdisplayshortskip}{0pt}

@@

\section{A Better Algorithm}
\subsection{Overview}
While the brute force algorithm is simple and correct, it
is far too slow for larger inputs.  Fortunately, there are multiple
alternative TSP algorithms with a non-polynomial runtime
that is better than $O(n!)$.  We will explore in detail the
\textit{Held--Karp} algorithm, an $O(2^n)$ solution that uses
dynamic programming \cite{heldkarp}.  Then we will look at it
to the MST approximation algorithm and discuss the advantages
of each approach.

\subsection{Held--Karp}
The Held--Karp algorithm is an $O(2^n)$ approach to solving the TSP\@.
It is not fast enough to solve TSP instances with a large number of
cities, but the solution for the instances it can solve is guarenteed
to be correct.  The algorithm works as follows:
\par

Suppose $D = (d_{ij})$ is distance
matrix where $d_{ij} = d_{ji}$, $n$ is the number of rows in $D$,
$S$ is a subset of $\{1, 2,\dots,n\}$, and $x\in S$.  We define $C(S, x)$
to be the distance of the optimal tour that starts at $x$, visits each
destination exactly once, and ends at $1$ \cite{heldkarp}.  The recurrence
for $C(S,x)$ is:

\begin{align*}
  C(S, x)=\begin{cases}
    \min\{k\in S-\{x\} : d_{kx} + C(S-\{x\}, k) \}  & \text{if } n>1 \\
    d_{1x}                                          & \text{if } n=1
  \end{cases}
\end{align*}

We obtain the optimal cost from a distance matrix $D$ of size $n$ by calling
$C(\{1,\dots,n\}, 1)$.  We can apply dynamic programming principles to this
problem by storing the cost for each call in a lookup table $H$.  In each
recurrence where $n>1$, if the tuple $(S, x)$ exists in $H$, we can return
$H[(S, x)]$.  Otherwise,  we will compute the cost and store it in $H$.
\par

This alone does not solve the original problem.  We still need to return 
the optimal tour, not just its distance.  We can find the optimal tour
by backtracing through $H$.  We define the optimal tour as $P$ and $B(v)$
as a recurrence that backtraces through $H$:

\begin{align*}
  B(v)=\begin{cases}
    B(H[v])     & \text{if } v \text{\ in\ } H \\
    \const{nil} & \text{otherwise} 
  \end{cases}
\end{align*}

We fill in the values of $P$ by calling $B((\{1,\dots,n\}, 1))$.  On each
recurrence where $v$ is in $H$, we append the second value from the tuple $v$
to $P$.  When $v$ is not in $H$ and $\const{nil}$ is returned from $B(v)$, we
append $1$ to $P$, resulting in the optimal tour.
\par

\subsection{Christofides}
Christofide's algorithm is an $O(n^3)$ approximation algorithm that guarentees
the cost of its solution will be strictly within $3/2$ of the optimal solution
cost.  The algorithm only works on a symmetric distance matrix that satisfies
the triangle inequality \cite{nicos}.  That is, a distance matrix $D=d_{ij}$
satisfies the triangle inequality if the shortest path to reach every $j$ from
$i$ involves traveling directly to $j$ from $i$, rather than through a path that
includes more than two verticies.
\par

The algorithm works as follows: Given a distance matrix $D=d_{ij}$,

% Appendicies
% ---
%\newpage
%\appendix
%\section{Appendices}
%\addcontentsline{toc}{section}{Appendices}
%\renewcommand{\thesubsection}{\Alph{subsection}}
%
%\subsection{Collaboration}
%  I collaborated with Grant Eaton and Luke Artnak.
%
%\newpage
%\subsection{Code Demonstration}
%  \inputminted[frame=lines,linenos=true]{python}{../src/brute_force.py}
%
\newpage
\bibliographystyle{apacite}
\bibliography{biblio}
